\papertitle

### 2013-01-17 manuscript.tex
Added MAST and NSF funding acknowledgements
Correct spelling of "ornithopter"

We introduce a new ornithopter MAV, the 13 gram H$^2$Bird. It 
features clap and fling wings, improves upon previous designs by 
utilizing a carbon fiber airframe, tail rotor, and elevator, and carries a 
2.8 gram payload. We augment the ornithopter's built-in gyroscope-based 
control with a lightweight ground station design, which has power and weight 
requirements appropriate for deployment on a millirobotic ground vehicle. The ground 
station provides heading estimates to the ornithoper by running a real-time 
motion tracking algorithm on a live video stream.
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ACM METADATA

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Robotic and Vision Platforms}
%----------------------------------------------------------------------------%
\subsection{H$^2$Bird Ornithoper}
% humphrey
To explore these cooperative control concepts, we designed a new 
flapping-wing MAV, known as the H$^2$Bird (Fig.~\ref{fig:h2bird}). Built 
around the Silverlit i-Bird RC flyer power train and clap-fling wings\footnote{\raggedright Silverlit Toys Manufactory Ltd.: i-Bird RC Flyer
\href{http://www.silverlit-flyingclub.com/wingsmaster/}
     {http://www.silverlit-flyingclub.com/wingsmaster/}}, 
the H$^2$Bird has a wingspan of 26.5 cm and a flight weight of 13 grams. A 
tail propeller and servo-controlled tail elevator provide control in the yaw 
and pitch axes (Fig.~\ref{fig:h2Bird_axes}). The on-board ImageProc 2.4\footnote{ImageProc 2.4: \\
\href{https://github.com/biomimetics/imageproc\_pcb}
     {https://github.com/biomimetics/imageproc\_pcb}} 
controller includes of a 40 MIPS microprocessor, 6 DOF IMU, 
IEEE 802.15.4 radio, VGA camera, and motor drivers, all powered by a 90 mAh 
lithium polymer battery. In routine flight, the H$^2$Bird averages 1.2 m/s 
ground speed, and operates for approximately 10 minutes.

@@

%----------------------------------------------------------------------------%
\subsection{Ornithoper Attitude Control}

@@

%............................................................................%
\subsubsection{Ornithoper Attitude Control}
\label{sec:flight_control}
We determined the turning speed and turning radius of the H$^2$Bird 
experimentally, by measuring the response of the system to a step input of a 
clockwise 90$^{\circ}$ turn. We measured the response of the system by 
recording the attitude estimates calculated by the ornithopter using its 
on-board gyroscope. We then used MATLAB\footnote{The MathWorks, Inc. Matlab:
\href{http://www.mathworks.com/products/matlab/}
     {http://www.mathworks.com/products/matlab/}} 
to fit a simple low-order model to the step response, as discussed in 
Sec.~\ref{sec:model_attitude}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}
The authors would like to thank Fernando Garcia Bermudez for his 
assistance with the Vicon motion capture system, Andrew Pullin for his 
help with robot photography, and the members of the Biomimetic 
Millisystems Laboratory and the EECS community at the University of 
California, Berkeley for their advice and support.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

We introduce a new ornithopter MAV, the 13 gram H$^2$Bird. It 
features clap and fling wings, improves upon previous designs by 
utilizing a carbon fiber airframe, tail rotor, and elevator, and carries a 
2.8 gram payload. We augment the ornithopter's built-in gyroscope-based 
control with a lightweight ground station design, which has power and weight 
requirements appropriate for deployment on a millirobotic ground vehicle. The ground 
station provides heading estimates to the ornithopter by running a real-time 
motion tracking algorithm on a live video stream.
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ACM METADATA

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Robotic and Vision Platforms}
%----------------------------------------------------------------------------%
\subsection{H$^2$Bird Ornithopter}
% humphrey
To explore these cooperative control concepts, we designed a new 
flapping-wing MAV, known as the H$^2$Bird (Fig.~\ref{fig:h2bird}). Built 
around the Silverlit i-Bird RC flyer power train and clap-fling wings\footnote{\raggedright Silverlit Toys Manufactory Ltd.: i-Bird RC Flyer
\href{http://www.silverlit-flyingclub.com/wingsmaster/}
     {http://www.silverlit-flyingclub.com/wingsmaster/}}, 
the H$^2$Bird has a wingspan of 26.5 cm and a flight weight of 13 grams. A 
tail propeller and servo-controlled tail elevator provide control in the yaw 
and pitch axes (Fig.~\ref{fig:h2Bird_axes}). The on-board ImageProc 2.4\footnote{ImageProc 2.4: \\
\href{https://github.com/biomimetics/imageproc\_pcb}
     {https://github.com/biomimetics/imageproc\_pcb}} 
controller includes of a 40 MIPS microprocessor, 6 DOF IMU, 
IEEE 802.15.4 radio, VGA camera, and motor drivers, all powered by a 90 mAh 
lithium polymer battery. In routine flight, the H$^2$Bird averages 1.2 m/s 
ground speed, and operates for approximately 10 minutes.

@@

%----------------------------------------------------------------------------%
\subsection{Ornithopter Attitude Control}

@@

%............................................................................%
\subsubsection{Ornithopter Attitude Control}
\label{sec:flight_control}
We determined the turning speed and turning radius of the H$^2$Bird 
experimentally, by measuring the response of the system to a step input of a 
clockwise 90$^{\circ}$ turn. We measured the response of the system by 
recording the attitude estimates calculated by the ornithopter using its 
on-board gyroscope. We then used MATLAB\footnote{The MathWorks, Inc. Matlab:
\href{http://www.mathworks.com/products/matlab/}
     {http://www.mathworks.com/products/matlab/}} 
to fit a simple low-order model to the step response, as discussed in 
Sec.~\ref{sec:model_attitude}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}
The authors would like to thank Fernando Garcia Bermudez for his 
assistance with the Vicon motion capture system, Andrew Pullin for his 
help with robot photography, and the members of the Biomimetic 
Millisystems Laboratory and the EECS community at the University of 
California, Berkeley for their advice and support. This material is based upon 
work supported by the U.S. Army Research Laboratory under the Micro Autonomous 
Systems and Technology Collaborative Technology Alliance, and the National 
Science Foundation under Grant No. IIS-0931463.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
